---
title: "Radio and Quarto"
description: |
  Music viz
date: 2024-04-23
---

April 23rd, 2024. Can we start the countdown to Gemini szn?

It's 6 AM here at the CJSW radio studio. We're broadcasting some Tom Skinner this morning. A brand new 20 minute live track from his forthcoming record, Voices of Bishara Live at "Mu".

I've wanted to do a little data/radio crossover for a bit so that's what I've focused on today. I have my full 4 year historical tracklist saved in a Google Sheet so why not see if I can read that into Quarto and play around with the data.

`googlesheets4` appears to be the Posit built R interface to Google Sheets so I'm going to give that a try. Fun fact - the 4 in `googlesheets4` references the version of the Sheets API. It's not because Jennifer Bryan (the developer of `googlesheets4`) liked the number 4. In my opinion, that would have been a good enough reason :)

Since my radio tracklist is saved in a private Google Sheet, authentication will be required. Looks like `googledrive` is needed to sort that out.

```{r}
# library(googledrive)
# library(googlesheets4)
# 
# drive_auth()
# gs4_auth(token = drive_token())
```

Okay authentication was a lot easier than I thought it would be. Two functions: `drive_auth()` and `gs4_auth()` is all it took, along with some browser clicks that lets the API access my Google Drive. We should be in business now.

```{r}
# tracklist <- read_sheet(
#   "https://docs.google.com/spreadsheets/d/1L_aDq6GCrTqJQqjgYCsWrv5l3QYM2z09rrs-11KnhvM/edit#gid=0",
#   "Tracklist",
#   skip = 1
# )
```
