---
title: "chattr"
description: |
  AI in RStudio
author: Michael Ellsworth
date: 2024-04-24
---

I came across [this blog post](https://blogs.rstudio.com/ai/posts/2024-04-04-chat-with-llms-using-chattr/) yesterday while scrolling through #rstats Mastodon and it inspired me to give `chattr` a try in RStudio. Based on the post, it looks like `chattr` simply allows for interaction with LLMs via a Shiny app in RStudio (or outside RStudio in a terminal).

Here is the install script:

```{r}
#| eval: false
# Install from GitHub
library(remotes)
remotes::install_github("mlverse/chattr")
```

In the "Getting started" section of the post, running the app is done via the `chattr_app()` function but I'm getting a "No model setup found" error. Running `?chattr_use` in the R Console pulls up some details to resolve this error. Looks like the easiest way to setup a model is to grab an OpenAI Personal Access Token (PAT). My preference would be to run a local LLM but maybe I can look into setting that up later.

After creating a PAT on the OpenAI website, I created an .Renviron file in my project directory and saved the token there under `OPENAI_API_KEY`. It looks like that's all it took to set-up the model. I can now run `chattr_app()`.

```{r}
library(chattr)
chattr::chattr_app()
```

This seems to open the Shiny App but sending a prompt to OpenAI doesn't return any response. After an unfortunate amount of time looking through Google, running `chattr::chattr("hello")` gives me an error message that might be a clue to resolving my problem:

```         
{error}Type:insufficient_quota
Message: You exceeded your current quota, please check your plan and billing details.
```

Based on my usage details on the OpenAI website, I should have plenty of quota to work with. Other folks online have reported a similar issue and have noted that it goes away after a few days. Perhaps I'll check back on this later.

This unsuccessful AI exercise reminds me that I need to read Molly White's article [AI isn't useless. But is it worth it?](https://www.citationneeded.news/ai-isnt-useless/). Unfortunately, I'm not sure this foray into `chattr` was worth much today!
