---
title: "5 Weeks - Day 11"
description: |
  YYC Weather
author: Michael Ellsworth
date: 2024-04-30
---

It's April 30th, 2024 and it snowed today. A lot. It snowed a lot. This inspired me to look at the Government of Canada's weather records for Calgary to see what data I could explore. This led me to [Almanac Averages & Extremes](https://climate.weather.gc.ca/climate_data/almanac_e.html?month=4&day=30&StationID=2205&period=1&searchMethod=contains&txtStationName=calgary). Let's take a look at the downloadable data!

```{r}
#| echo: false
#| warnings: false
library(readr)
library(dplyr)
library(janitor)
library(lubridate)

almanac_url <- "https://climate.weather.gc.ca/climate_data/bulk_data_e.html?format=csv&stationID=2205&Year=2008&Month=4&Day=30&time=&timeframe=4&submit=Download+"

df <- read_csv(almanac_url) %>% janitor::clean_names()
df
```

Looks like we have some weather records on our hands! Let's take a look at April 30th to see what type of extremes we have dealt with on this day in the past.

```{r}
df %>%
  filter(month == 4, day == 30) %>%
  select(
    matches("average"), 
    matches("highest"), 
    matches("lowest"), 
    matches("greatest"),
    matches("most")
  )
```

I'm curious to compare these with today's conditions but we don't have that data yet. Let's come back to April 30th, perhaps tomorrow. In the meantime, let's take a look at April 29th! We had a bit of rain and snow yesterday so let's compare that to some historical records.

```{r}
df_2024_04_29 <- df %>%
  filter(month == 4, day == 29) %>%
  select(
    matches("average"), 
    matches("highest"), 
    matches("lowest"), 
    matches("greatest"),
    matches("most")
  )

daily_report_url <- "https://climate.weather.gc.ca/climate_data/bulk_data_e.html?format=csv&stationID=50430&Year=2024&Month=4&Day=1&time=&timeframe=2&submit=Download+Data"

df_2024 <- read_csv(daily_report_url) %>%
  janitor::clean_names() %>%
  mutate(date_time = lubridate::ymd(date_time))
  
```

```{r}

```
